---
title: "Day 4 Activity 1: Example Solution"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = NA)
library(ggplot2)
```

## Activity 1: What Is Wrong With This Code?

This example solution explains the main reliability problems in the current
`starter.R` workflow before showing one corrected version.

### Potential Issues In The Starter Code

1. `compute_iou()` does not validate boxes or missing coordinates.
The mock CSV includes invalid rows and one row with missing coordinates, so
malformed input can quietly flow through the calculation instead of failing
clearly.

2. `compute_iou()` does not clamp negative overlap widths and heights at zero.
That means non-overlapping boxes can still produce nonsense IoU values.

3. `build_agreement_curve()` uses `na.rm = TRUE`.
Once bad rows are converted to `NA`, those rows disappear from the denominator
and the agreement rate is artificially inflated.

### Example Solution

```{r}
validate_box <- function(box) {
  if (length(box) != 4) {
    stop("box must contain x1, y1, x2, y2")
  }

  if (any(is.na(box))) {
    stop("box cannot contain missing coordinates")
  }

  if (box[3] <= box[1] || box[4] <= box[2]) {
    stop("box coordinates must satisfy x2 > x1 and y2 > y1")
  }
}

compute_iou <- function(box_a, box_b) {
  validate_box(box_a)
  validate_box(box_b)

  x_left <- max(box_a[1], box_b[1])
  y_top <- max(box_a[2], box_b[2])
  x_right <- min(box_a[3], box_b[3])
  y_bottom <- min(box_a[4], box_b[4])

  inter_width <- max(0, x_right - x_left)
  inter_height <- max(0, y_bottom - y_top)
  inter_area <- inter_width * inter_height

  area_a <- (box_a[3] - box_a[1]) * (box_a[4] - box_a[2])
  area_b <- (box_b[3] - box_b[1]) * (box_b[4] - box_b[2])
  union_area <- area_a + area_b - inter_area

  if (union_area <= 0) {
    stop("union area must be positive")
  }

  inter_area / union_area
}

compute_iou_safe <- function(box_a, box_b) {
  tryCatch(
    compute_iou(box_a, box_b),
    error = function(e) NA_real_
  )
}

add_iou_all_frames <- function(box_pairs) {
  box_pairs <- box_pairs

  for (row in 1:nrow(box_pairs)) {
    box_a <- c(
      box_pairs[row, "box_a_x1"],
      box_pairs[row, "box_a_y1"],
      box_pairs[row, "box_a_x2"],
      box_pairs[row, "box_a_y2"]
    )
    box_b <- c(
      box_pairs[row, "box_b_x1"],
      box_pairs[row, "box_b_y1"],
      box_pairs[row, "box_b_x2"],
      box_pairs[row, "box_b_y2"]
    )

    box_pairs[row, "iou"] <- compute_iou_safe(box_a, box_b)
  }

  box_pairs
}

build_agreement_curve <- function(box_pairs) {
  thresholds <- seq(0, 1, by = 0.05)
  data.frame(
    iou_threshold = thresholds,
    percent_agreement = sapply(
      thresholds,
      function(threshold) {
        100 * mean(ifelse(is.na(box_pairs$iou), FALSE, box_pairs$iou >= threshold))
      }
    )
  )
}
```

### Example Checks

```{r}
csv_path <- file.path("modules", "day-04-reliable", "activities", "activity-1", "mock_box_pairs.csv")
if (!file.exists(csv_path)) {
  csv_path <- "mock_box_pairs.csv"
}

box_pairs <- read.csv(csv_path, check.names = FALSE)
box_pairs_w_iou <- add_iou_all_frames(box_pairs)
curve_info <- build_agreement_curve(box_pairs_w_iou)

stopifnot(all(box_pairs_w_iou$iou[box_pairs_w_iou$scenario == "touching_edge"] == 0))
stopifnot(sum(is.na(box_pairs_w_iou$iou)) == 21)
stopifnot(sum(is.na(box_pairs_w_iou$iou[box_pairs_w_iou$scenario == "invalid_box"])) == 20)
stopifnot(sum(is.na(box_pairs_w_iou$iou[box_pairs_w_iou$scenario == "no_overlap"])) == 1)
stopifnot(all(box_pairs_w_iou$iou[box_pairs_w_iou$scenario == "normal_overlap"] > 0))
stopifnot(curve_info$percent_agreement[curve_info$iou_threshold == 0] == 89.5)

ggplot(curve_info, aes(x = iou_threshold, y = percent_agreement)) +
  geom_line() +
  geom_point() +
  labs(x = "IoU Threshold", y = "% Agreement") +
  theme_bw()
```
