---
title: "Differential Gene Expression Analysis"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = NA)
```

# Differential Gene Expression Analysis

Author: Tuan Nguyen

This notebook identifies significantly differentially expressed genes between
the Treatment and Control groups.

## Import The Reusable Function

Keep the notebook focused on parameters, outputs, and interpretation. The core
analysis should stay in the imported function.

```{r}
library(dplyr)
library(ggplot2)
source("refactor_function.R")
```

## Set Notebook Parameters

Keep the values that vary near the top of the notebook so the next person can
rerun it quickly.

```{r}
expression_file <- "simulated_expression_matrix.csv"
alpha <- 0.05
```

**Alternative Option**: Keep Configurations Separately

If you rerun the notebook often, you can move the changing values into a small
YAML file instead of editing the notebook itself each time.

```{r}
config_path <- "config.yaml"
config <- yaml::read_yaml(config_path)

expression_file <- if (!is.null(config$expression_file)) config$expression_file else "simulated_expression_matrix.csv"
alpha <- if (!is.null(config$alpha)) config$alpha else 0.05
```

**Recommendation**: Print The Chosen Configurations

```{r}
cat("Using expression file:", expression_file, "\n")
cat("Using alpha:", alpha, "\n")
```

## Run The Analysis And View Results

The notebook wrapper should call the reusable function and keep the workflow
readable. Describe what each code chunk does.

### Perform Statistical Testing To Identify Significant Genes

```{r}
df <- read.csv(expression_file, row.names = 1)
significant_de_genes <- find_de_genes(df, alpha = alpha)
```

```{r}
cat("Number of significant genes:", nrow(significant_de_genes), "\n")
cat("Top significant genes:\n")
print(head(significant_de_genes))
```

## Quick Visualization

```{r}
significant_de_genes$log10_padj <- -log10(as.numeric(significant_de_genes$padj))

ggplot(significant_de_genes, aes(x = fc, y = log10_padj)) +
  geom_point() +
  geom_text(aes(label = gene), vjust = -0.5, size = 3) +
  labs(x = "Fold Change",
       y = "-log10 (adjusted p-value)") +
  theme_bw()
```
