---
title: "Gene Correlation Analysis"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = NA)
library(dplyr)
library(tidyr)
library(tibble)
library(ggplot2)
```

# Gene Correlation Analysis

This notebook identifies and plots the expression of genes highly correlated
with a target gene across all tissues.

### Import the reusable functions

```{r}
# TODO: Import the helper functions from activity 1 and others (if necessary)
# You can use either source(), or copy-paste the function in a separate code chunk.
```

### Set parameters for the analysis

The number of parameters depends on how many arguments you put in your helper
functions.

You can choose to type them in differently each time, or have a separate yaml
file.

```{r}
data_dir <- "../activity-1/mock_data_3_tissues"
tissue_files <- c(
  Adipose = "Adipose.csv",
  Liver = "Liver.csv",
  Muscle = "Muscle.csv"
)

# TODO:
# Add the necessary parameters for the analysis.
# Make sure the parameters are known when the report is compiled.
```

### Run the actual analysis

The repeated code logic for every iteration. You should be able to copy-paste
it from the solution to Activity 1.

#### Identify the correlated genes

```{r}
# TODO: Implement the analysis that should be repeated across all iterations.
```

#### Plot their expression in three tissues

```{r}
# TODO: Visualize the expression for selected genes for each tissue.
```
