---
title: "Reusable Analysis Notebook Template"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = NA)
```

## Goal

Use this notebook template to wrap your refactored analysis in a small,
parameterized workflow.

## Choose The Values That Vary

Fill in the values you want to make easy to change between runs.

```{r}
data_path <- "TODO"
alpha <- 0.05
```

## Import Your Refactored Function

Replace the path below with the file that contains your cleaned function from
Activity 1.

```{r}
source("TODO")
```

## Wrap The Reusable Step

Keep this wrapper smaller than the analysis function it calls.

```{r}
run_de_analysis <- function(data_path, alpha = 0.05) {
  expression_matrix <- read.csv(data_path, row.names = 1)

  # TODO:
  # Call your imported function here.
  # Return the result.
}
```

## Run The Workflow

```{r}
res <- run_de_analysis(data_path, alpha = alpha)

cat("Alpha:", alpha, "\n")
cat("Number of significant genes:", nrow(res), "\n")
print(head(res))
```

## Reflection

Add 1-2 short notes:

- Which parameter improved reuse the most?
- Which value did you decide to keep fixed?
