---
title: "Day 4 - Demo 1: Silent Break"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = NA)
library(dplyr)
library(ggplot2)
library(tidyr)
library(zoo)
```

## Day 4 - Demo 1

Two silent-break examples with plausible but wrong behavior.

This version uses native R code.

#### 1a. Boxplot

You have numeric measurements from three different conditions, both sexes.

You want to visualize them in a boxplot. What's wrong with this code?

```{r}
plot_boxplot <- function(data, x, y, fill) {
  ggplot(data, aes(x = .data[[x]], y = .data[[y]], fill = .data[[fill]])) +
    geom_boxplot(position = position_dodge(width = 0.75)) +
    geom_point(aes(group = .data[[fill]]),
               position = position_jitterdodge(jitter.width = 0.15, dodge.width = 0.75)) +
    labs(x = x, y = y, fill = fill, color = fill) +
    theme_minimal()
}
```

Here is a helper function to generate mock data.

```{r}
generate_mock_data <- function(num_samples = 500, seed = NULL) {
  if (!is.null(seed)) {
    set.seed(seed)
  }

  tibble(
    Sample_ID = sprintf("Sample_%d", seq_len(num_samples)),
    Condition = sample(c("A", "B", "C"), size = num_samples, replace = TRUE),
    Sex = sample(c("Male", "Female"), size = num_samples, replace = TRUE),
    Measurement = rnorm(num_samples, mean = 0, sd = 1)
  )
}
```

```{r}
demo1a_mock_data <- generate_mock_data(num_samples = 500, seed = 1)
plot_boxplot(demo1a_mock_data, x = "Condition", y = "Measurement", fill = "Sex")
```

#### 1b. Moving average

You have noisy time-series data for many individuals. Each row is a sample, and each column is the minute from which the measurement was recorded. 

Because the data is quite noisy, I smooth the data using a moving average. Then I plot the trend over time aggregated for each group.

What can potentially be wrong with the code?

```{r}
#' Plot smoothed time-series data by condition and sex
#'
#' @description
#' Plot smoothed time-series data by condition (color) and sex (style).
#'
#' @param data A data.frame containing 'Condition', 'Sex', and numeric time columns (e.g., '1', '2', ...).
#' @param window Integer. Rolling window for smoothing across time columns. Default is 5.
#' @param figsize Numeric vector of length 2. Figure size. Default is c(10, 6).
#'
#' @return A \code{matplotlib.figure.Figure}-like object representing the plot of time-smoothed measurements.
plot_smoothed_measurements <- function(data, window = 5, figsize = c(10, 6)) {
  # Identify time-series columns (numeric names like "1","2",...)
  measurement_cols <- colnames(data)[grepl("^\\d+$", colnames(data))]

  # Smooth across time for each sample
  smoothed_data <- data %>%
    mutate(across(all_of(measurement_cols), 
           ~ rollapply(.x, width = window, FUN = mean, align = "center", partial = TRUE))) %>%
    as.data.frame()

  # Convert to long format
  long_df <- smoothed_data %>%
    pivot_longer(cols = all_of(measurement_cols), 
                 names_to = "Time",
                 values_to = "Value") %>%
    mutate(Time = as.numeric(Time))

  long_summary <- long_df %>%
    dplyr::group_by(Condition, Sex, Time) %>%
    dplyr::summarise(Value = mean(Value, na.rm = TRUE), .groups = "drop")
  
  p <- ggplot(long_summary, aes(x = Time, y = Value, color = Condition, linetype = Sex)) +
    geom_line() +
    labs(title = "Smoothed Measurements", x = "Time", y = "Value") +
    theme_bw()

  return(p)
}
```

Here is a helper function to generate mock data and test the function.

```{r}
generate_mock_time_series <- function(num_samples = 80, num_timepoints = 90, seed = NULL) {
  if (!is.null(seed)) set.seed(seed)

  # Time series data with explicit column names
  time_cols <- as.character(0:(num_timepoints - 1))
  time_series_data <- as.data.frame(matrix(
    runif(num_samples * num_timepoints),
    nrow = num_samples,
    ncol = num_timepoints
  ))
  colnames(time_series_data) <- time_cols

  # Metadata
  metadata <- data.frame(
    Sample_ID = paste0("Sample_", seq_len(num_samples)),
    Condition = sample(c("A", "B"), num_samples, replace = TRUE),
    Sex = sample(c("Male", "Female"), num_samples, replace = TRUE)
  )

  # Combine
  data <- cbind(metadata, time_series_data)

  return(data)
}
```

```{r}
demo1b_mock_data <- generate_mock_time_series()
plot_smoothed_measurements(demo1b_mock_data, window = 5)
```

### Example improved code

For Demo 1a, one reliability issue is the artificial duplication of outliers in
plotting: one point comes from the boxplot itself, and another comes from the
overlaid points.

The fix is to hide the outliers in the boxplot layer.

```{r}
plot_boxplot_improved <- function(data, x, y, fill) {
  ggplot(data, aes(x = .data[[x]], y = .data[[y]], fill = .data[[fill]])) +
    geom_boxplot(position = position_dodge(width = 0.75), outlier.shape = NA) +
    geom_point(aes(group = .data[[fill]]), position = position_jitterdodge(jitter.width = 0.15, dodge.width = 0.75)) +
    labs(x = x, y = y, fill = fill, color = fill) +
    theme_minimal()
}
```

For Demo 1b, there are several reliability issues with the rolling average.

- The rolling average is going along the wrong axis, so the smoothing happens
  across samples instead of across time.
- It assumes the time columns are already in the correct increasing order.
- It assumes the time points are evenly spaced.

We can identify more problems, but this example only fixes those three main
issues for simplicity.

This function is also doing multiple things at once, with separate smoothing
and plotting responsibilities. It is usually clearer to break that into smaller
functions.

```{r}
get_time_columns <- function(data) {
  ### Identify time columns in the DataFrame.
  
  # Identify columns that are purely digits
  time_cols <- names(data)[grepl("^\\d+$", names(data))]
  # Ensure correct temporal order
  time_cols <- time_cols[order(as.integer(time_cols))]

  # Confirm evenly spaced timepoints
  timepoints <- as.integer(time_cols)
  if (length(timepoints) > 1) {
    diffs <- diff(timepoints)
    if (!all(diffs == diffs[1])) {
      stop("Timepoints are not evenly spaced.", call. = FALSE)
    }
  }

  return(time_cols)
}

smooth_time_series <- function(data, window = 5) {
  ### Apply rolling mean smoothing to time-series data.
  measurement_cols <- get_time_columns(data)
  
  # Smooth across time for each sample
  smoothed_matrix <- data %>%
    dplyr::select(all_of(measurement_cols)) %>%
    t() %>% as.data.frame() %>%
    mutate(across(everything(), 
           ~ rollapply(.x, width = window, FUN = mean, align = "center", partial = TRUE))) %>%
    t()

  smoothed_data <- data
  smoothed_data[measurement_cols] <- as.data.frame(smoothed_matrix)

  smoothed_data %>%
    pivot_longer(
      cols = all_of(measurement_cols),
      names_to = "Time",
      values_to = "Value"
    ) %>%
    mutate(Time = as.numeric(Time))
}

plot_smoothed_measurements_improved <- function(data, window = 5, figsize = c(10, 6)) {
  smoothed_data_long <- smooth_time_series(data, window = window)

  long_summary <- smoothed_data_long %>%
    group_by(Condition, Sex, Time) %>%
    summarize(Value = mean(Value, na.rm = TRUE), .groups = "drop")

  ggplot(long_summary, aes(x = Time, y = Value, color = Condition, linetype = Sex)) +
    geom_line() +
    labs(title = "Smoothed Measurements", x = "Time", y = "Value") +
    theme_bw()
}
```
